---
title: "The Human Lung Cell Atlas (10X)"
output:
  html_document: default
  html_notebook: default
---
# All Package versions used


```{r, fig.width=10, fig.height=10}
installed.packages()[,'Version']

```

# Specify the tissue of interest, run the boilerplate code which sets up the functions and environment, load the tissue object.


```{r, fig.width=10, fig.height=10}
datasets = c("GSM3489182_Donor_01", "GSM3489187_Donor_03", "GSM3489189_Donor_04", "GSM3489193_Donor_06", "GSM3489197_Donor_08")
library(here)
source(here("boilerplate.R"))
tiss10x <- load_tissue_droplet(datasets)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x, pc.use = 1:9, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x, pc.use = 10:18, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
```

```{r, fig.width=10, fig.height=10}
PCElbowPlot(object = tiss10x)
```

```{r, fig.width=10, fig.height=10}
# Set number of principal components.
pcs = c(1:5,8,10,11,14,15,16,17,18)
```

```{r, fig.width=10, fig.height=10}
# Set resolution
res.used <- 2

tiss10x <- FindClusters(object = tiss10x, reduction.type = "pca", dims.use = pcs,
    resolution = res.used, print.output = 0, save.SNN = TRUE)
```

```{r, fig.width=10, fig.height=10}
tiss10x <- RunTSNE(object = tiss10x, dims.use = pcs, seed.use = 10, perplexity=30)
```


```{r, fig.width=10, fig.height=10}
# note that you can set do.label=T to help label individual clusters
TSNEPlot(object = tiss10x, do.label = T, no.legend = T, label.size = 10)
```

```{r, fig.width=10, fig.height=10}
# Batch and sample effects
TSNEPlot(object = tiss10x, do.return = TRUE, group.by = "orig.ident", no.legend = T)
```

```{r, fig.height=10, fig.width=10}
genes_to_check = c('CLDN5', 'EPCAM', 'COL1A2', 'PTPRC')
FeaturePlot(tiss10x, genes_to_check, pt.size = 1, nCol = 2, cols.use = my.cols, dark.theme = TRUE)
```

```{r, fig.height=10, fig.width=10}
FeaturePlot(tiss10x, 'nGene', pt.size = 1, nCol = 1, cols.use = my.cols, dark.theme = TRUE)
```

# Subclustering normal stromal cells

```{r, fig.width=10, fig.height=10}
cells_to_use <- WhichCells(tiss10x, ident = 24)
cells_to_use <- intersect(cells_to_use, names(which(tiss10x@data['SFTPA1',] < 3)))
tiss10x.stromal = SubsetData(tiss10x, cells.use = cells_to_use)
```

```{r, fig.width=10, fig.height=10}
tiss10x.stromal <- tiss10x.stromal %>% ScaleData() %>%
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE) %>% ProjectPCA(do.print = FALSE)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x.stromal, pc.use = 1:9, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.stromal, pc.use = 10:18, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCElbowPlot(tiss10x.stromal)
```


```{r, fig.width=10, fig.height=10}
normal.n.pcs = 4
normal.res.use = 1
tiss10x.stromal <- tiss10x.stromal %>% FindClusters(reduction.type = "pca", dims.use = 1:normal.n.pcs,
    resolution = normal.res.use, print.output = 0, save.SNN = TRUE, plot.SNN = TRUE, k.param = 10) %>%
    RunTSNE(dims.use = 1:normal.n.pcs, seed.use = 10, perplexity=20)

TSNEPlot(object = tiss10x.stromal, do.label = T, pt.size = 3, label.size = 10)

```

```{r, fig.width=10, fig.height=10}
# Batch and sample effects
TSNEPlot(object = tiss10x.stromal, do.return = TRUE, group.by = "orig.ident", pt.size = 4, label.size = 10)
```

```{r, fig.height=50, fig.width=10}

genes_to_check = c('CLDN5', 'EPCAM', 'COL1A2', 'PTPRC', 'SPINT2', 'SFRP2', 'FGFR4', 'ITGA8', 'GPC3', 'ASPN', 'FGF18', 'HIGD1B', 'KCNK3', 'ACTA2', 'CNN1', 'TAGLN', 'DES', 'KCNA5', 'ADAMTS4', 'PDGFRA', 'PDGFRB')

FeaturePlot(tiss10x.stromal, genes_to_check, pt.size = 4, nCol = 2, cols.use = my.cols, dark.theme = TRUE)

```

```{r, fig.width=10, fig.height=10}
tiss10x.stromal.markers <- FindAllMarkers(tiss10x.stromal, min.pct = 0.25, thresh.use = 0.25, only.pos = T)

```


```{r, fig.height=80, fig.width=10}
genes_to_check <- tiss10x.stromal.markers %>% group_by(cluster) %>% top_n(-4, p_val_adj)
tiss10x.stromal.markers %>% group_by(cluster) %>% top_n(-10, p_val_adj)
FeaturePlot(tiss10x.stromal, genes_to_check$gene, pt.size = 3, nCol = 2, cols.use = my.cols, dark.theme = TRUE)
```

## Assigning cell type identity to clusters


```{r, fig.width=10, fig.height=10}
#SERPINF1+ PI16+ FBLN1+ SCARA5+ SLC38A5- -> Adventitial Fibroblast
#ACTA2+ TAGLN+ CNN1+ MYH11+ -> Smooth Muscle
#SLC38A5+ GDF10+ GPC3+ ITGA8+ -> Alveolar Fibroblast


# stash current cluster IDs
tiss10x.stromal <- StashIdent(object = tiss10x.stromal, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- 0:max(as.numeric(levels(tiss10x.stromal@ident)))
free_annotation <- c("Alveolar Fibroblast",
                     "Alveolar Fibroblast",
                     "Smooth Muscle",
                     "Adventitial Fibroblast")

tiss10x.stromal@meta.data[,'free_annotation'] <- NA
tiss10x.stromal@meta.data[,'free_annotation'] <- as.character(plyr::mapvalues(x = tiss10x.stromal@ident, from = cluster.ids, to = free_annotation))
```

# Subclustering normal epithelial cells

```{r, fig.width=10, fig.height=10}
tiss10x.epi = SubsetData(tiss10x, ident.use = c(2,5,6,3,10,15,20,23,14,4,13))
```

```{r, fig.width=10, fig.height=10}
tiss10x.epi <- tiss10x.epi %>% ScaleData() %>%
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE) %>% ProjectPCA(do.print = FALSE)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x.epi, pc.use = 1:9, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.epi, pc.use = 10:18, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCElbowPlot(tiss10x.epi)
```


```{r, fig.width=10, fig.height=10}
normal.pcs = c(1,3,4,5,7,9)
normal.res.use = 1
tiss10x.epi <- tiss10x.epi %>% FindClusters(reduction.type = "pca", dims.use = normal.pcs,
    resolution = normal.res.use, print.output = 0, save.SNN = TRUE, plot.SNN = TRUE, force.recalc = TRUE, prune.SNN) %>%
    RunTSNE(dims.use = normal.pcs, seed.use = 10, perplexity=30)

TSNEPlot(object = tiss10x.epi, do.label = T, pt.size = 3, label.size = 10)

```

```{r, fig.width=10, fig.height=10}
# Batch and sample effects
TSNEPlot(object = tiss10x.epi, do.return = TRUE, group.by = "orig.ident")
```

```{r, fig.height=50, fig.width=10}

genes_to_check = c('CLDN5', 'EPCAM', 'COL1A2', 'PTPRC', 'MUC5AC', 'MUC5B', 'SFTPC', 'ETV5', 'PDPN', 'AGER', 'FOXJ1', 'TUBB4B', 'SCGB1A1', 'SCGB3A2', 'WIF1', 'HHIP', 'SERPINA1', 'KRT5', 'TP53', 'CALCA')

FeaturePlot(tiss10x.epi, genes_to_check, pt.size = 1, nCol = 2, cols.use = my.cols, dark.theme = TRUE)

```

```{r, fig.width=10, fig.height=10}
tiss10x.epi.markers <- FindAllMarkers(tiss10x.epi, min.pct = 0.25, thresh.use = 0.25, only.pos = T)

```


```{r, fig.height=60, fig.width=10}
genes_to_check <- tiss10x.epi.markers %>% group_by(cluster) %>% top_n(-4, p_val_adj)
tiss10x.epi.markers %>% group_by(cluster) %>% top_n(-10, p_val_adj)
FeaturePlot(tiss10x.epi, genes_to_check$gene, pt.size = 1, nCol = 2, cols.use = my.cols)
```

## Assigning cell type identity to clusters

```{r, fig.width=10, fig.height=10}
# At a coarse level, we can use canonical markers to match the unbiased clustering to known cell types:
# Relying on canonical markers to call populations (a few representative markers are shown). All epithelial cells are CD45- CD31- Col1A1-
# SFP*+ ETV5+ MUC1+ -> Alveolar Epithelial Type 2 Cell
# HOPX+ PDPN+ CLIC5+ SPC*- -> Alveolar Epithelial Type 1 Cell
# SCGB1A1+ SCGB3A2+ SP*- FOXJ1- -> Club Cells
# FOXJ1+ TUBB4+ -> Ciliated Cell
# KRT5+ KRT14+ -> Basal Cells (clustered again in Round 2)


# stash current cluster IDs
tiss10x.epi <- StashIdent(object = tiss10x.epi, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- 0:max(as.numeric(levels(tiss10x.epi@ident)))
free_annotation <- c('Alveolar Epithelial Type 2',
                     'Alveolar Epithelial Type 2',
                     'Alveolar Epithelial Type 2',
                     'Alveolar Epithelial Type 2',
                     'Alveolar Epithelial Type 2',
                     'Alveolar Epithelial Type 2',
                     'Alveolar Epithelial Type 2',
                     'Alveolar Epithelial Type 2',
                     NA,
                     'Alveolar Epithelial Type 1',
                     'Alveolar Epithelial Type 2',
                     'Alveolar Epithelial Type 2',
                     'Alveolar Epithelial Type 2',
                     NA)

tiss10x.epi@meta.data[,'free_annotation'] <- NA
tiss10x.epi@meta.data[,'free_annotation'] <- as.character(plyr::mapvalues(x = tiss10x.epi@ident, from = cluster.ids, to = free_annotation))
```



# Subclustering normal epithelial cells (Round 2, non-AT2)

```{r, fig.width=10, fig.height=10}
tiss10x.epi.r2 = SubsetData(tiss10x.epi, ident.use = c(8,13))
```

```{r, fig.width=10, fig.height=10}
tiss10x.epi.r2 <- tiss10x.epi.r2 %>% ScaleData() %>%
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE) %>% ProjectPCA(do.print = FALSE)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x.epi.r2, pc.use = 1:9, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.epi.r2, pc.use = 10:18, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCElbowPlot(tiss10x.epi.r2)
```


```{r, fig.width=10, fig.height=10}
normal.pcs = c(1,2,3,5)
normal.res.use = 1
tiss10x.epi.r2 <- tiss10x.epi.r2 %>% FindClusters(reduction.type = "pca", dims.use = normal.pcs,
    resolution = normal.res.use, print.output = 0, save.SNN = TRUE, force.recalc = TRUE, plot.SNN = TRUE, k.param = 10) %>%
    RunTSNE(dims.use = normal.pcs, seed.use = 10, perplexity=50)

TSNEPlot(object = tiss10x.epi.r2, do.label = T, pt.size = 3, label.size = 10)

```

```{r, fig.width=10, fig.height=10}
# Batch and sample effects
TSNEPlot(object = tiss10x.epi.r2, do.return = TRUE, group.by = "orig.ident", no.legend = T)
```

```{r, fig.height=50, fig.width=10}

genes_to_check = c('CLDN5', 'EPCAM', 'COL1A2', 'PTPRC', 'MUC5AC', 'MUC5B', 'SFTPC', 'ETV5', 'PDPN', 'AGER', 'FOXJ1', 'TUBB4B', 'SCGB1A1', 'SCGB3A2', 'WIF1', 'HHIP', 'SERPINA1', 'KRT5', 'TP53', 'CALCA', 'MKI67')

FeaturePlot(tiss10x.epi.r2, genes_to_check, pt.size = 2, nCol = 2, cols.use = my.cols, dark.theme = TRUE)

```

```{r, fig.width=10, fig.height=10}
tiss10x.epi.r2.markers <- FindAllMarkers(tiss10x.epi.r2, min.pct = 0.25, thresh.use = 0.25, only.pos = T)

```


```{r, fig.height=30, fig.width=10}
genes_to_check <- tiss10x.epi.r2.markers %>% group_by(cluster) %>% top_n(-4, p_val_adj)
tiss10x.epi.r2.markers %>% group_by(cluster) %>% top_n(-10, p_val_adj)
FeaturePlot(tiss10x.epi.r2, genes_to_check$gene, pt.size = 1, nCol = 2, cols.use = my.cols)
```

## Assigning cell type identity to clusters


```{r, fig.width=10, fig.height=10}
# stash current cluster IDs
tiss10x.epi.r2 <- StashIdent(object = tiss10x.epi.r2, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- 0:max(as.numeric(levels(tiss10x.epi.r2@ident)))
free_annotation <- c("Alveolar Epithelial Type 2",
                     "Alveolar Epithelial Type 2",
                     NA,
                     "Club",
                     "Alveolar Epithelial Type 2",
                     "Ciliated",
                     "Alveolar Epithelial Type 2",
                     "Alveolar Epithelial Type 2",
                     "Club",
                     NA,
                     "Alveolar Epithelial Type 2",
                     "Ciliated",
                     "Alveolar Epithelial Type 2",
                     "Alveolar Epithelial Type 2",
                     "Club")

tiss10x.epi.r2@meta.data[,'free_annotation'] <- NA
tiss10x.epi.r2@meta.data[,'free_annotation'] <- as.character(plyr::mapvalues(x = tiss10x.epi.r2@ident, from = cluster.ids, to = free_annotation))
```


# Subclustering normal endothelial cells

```{r, fig.width=10, fig.height=10}
cells_to_use = WhichCells(tiss10x, ident = c(17))
tiss10x.endo = SubsetData(tiss10x, cells.use = cells_to_use)
```


```{r, fig.width=10, fig.height=10}
tiss10x.endo <- tiss10x.endo %>% ScaleData() %>%
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE) %>% ProjectPCA(do.print = FALSE)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x.endo, pc.use = 1:9, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.endo, pc.use = 10:18, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCElbowPlot(tiss10x.endo)
```


```{r, fig.width=10, fig.height=10}
normal.n.pcs = 8
normal.res.use = 1
tiss10x.endo <- tiss10x.endo %>% FindClusters(reduction.type = "pca", dims.use = 1:normal.n.pcs,
    resolution = normal.res.use, print.output = 0, save.SNN = TRUE, force.recalc = TRUE) %>%
    RunTSNE(dims.use = 1:normal.n.pcs, seed.use = 10, perplexity=30)

TSNEPlot(object = tiss10x.endo, do.label = T, pt.size = 3, label.size = 4)

```

```{r, fig.width=10, fig.height=10}
# Batch and sample effects
TSNEPlot(object = tiss10x.endo, do.return = TRUE, group.by = "orig.ident")
```

```{r, fig.height=45, fig.width=10}

genes_to_check = c('CLDN5', 'EPCAM', 'COL1A2', 'PTPRC', 'DKK2', 'GJA5', 'ACKR1', 'PTGIS', 'EDNRB', 'CA4', 'IL7R', 'MYC', 'SPRY1', 'SOX17', 'PROX1', 'CCL21')

FeaturePlot(tiss10x.endo, genes_to_check, pt.size = 2, nCol = 2, cols.use = my.cols, dark.theme = TRUE)

```

```{r, fig.width=10, fig.height=10}
tiss10x.endo.markers <- FindAllMarkers(tiss10x.endo, min.pct = 0.25, thresh.use = 0.25, only.pos = T)

```


```{r, fig.height=80, fig.width=10}
genes_to_check <- tiss10x.endo.markers %>% group_by(cluster) %>% top_n(-4, p_val_adj)
tiss10x.endo.markers %>% group_by(cluster) %>% top_n(-10, p_val_adj)
FeaturePlot(tiss10x.endo, genes_to_check$gene, pt.size = 1, nCol = 2, cols.use = my.cols)
```

## Assigning cell type identity to clusters


```{r, fig.width=10, fig.height=10}
# stash current cluster IDs
tiss10x.endo <- StashIdent(object = tiss10x.endo, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- 0:max(as.numeric(levels(tiss10x.endo@ident)))
free_annotation <- c(NA,
                     'Artery',
                     'Lymphatic',
                     NA,
                     NA,
                     NA,
                     NA)

tiss10x.endo@meta.data[,'free_annotation'] <- NA
tiss10x.endo@meta.data[,'free_annotation'] <- as.character(plyr::mapvalues(x = tiss10x.endo@ident, from = cluster.ids, to = free_annotation))
```


# Subclustering endothelial cells (Round 2)

```{r, fig.width=10, fig.height=10}
cells_to_use = WhichCells(tiss10x.endo, ident = c(0,3,6))
tiss10x.endo.r2 = SubsetData(tiss10x.endo, cells.use = cells_to_use)
```


```{r, fig.width=10, fig.height=10}
tiss10x.endo.r2 <- tiss10x.endo.r2 %>% ScaleData() %>%
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE) %>% ProjectPCA(do.print = FALSE)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x.endo.r2, pc.use = 1:9, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.endo.r2, pc.use = 10:18, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCElbowPlot(tiss10x.endo.r2)
```


```{r, fig.width=10, fig.height=10}
normal.n.pcs = 3
normal.res.use = 1
tiss10x.endo.r2 <- tiss10x.endo.r2 %>% FindClusters(reduction.type = "pca", dims.use = 1:normal.n.pcs,
    resolution = normal.res.use, print.output = 0, save.SNN = TRUE, plot.SNN = TRUE, force.recalc = TRUE, k.param = 15) %>%
    RunTSNE(dims.use = 1:normal.n.pcs, seed.use = 10, perplexity=30)

TSNEPlot(object = tiss10x.endo.r2, do.label = T, pt.size = 3, label.size = 4)

```

```{r, fig.width=10, fig.height=10}
# Batch and sample effects
TSNEPlot(object = tiss10x.endo.r2, do.return = TRUE, group.by = "orig.ident")
```

```{r, fig.height=45, fig.width=10}

genes_to_check = c('CLDN5', 'EPCAM', 'COL1A2', 'PTPRC', 'DKK2', 'GJA5', 'ACKR1', 'PTGIS', 'EDNRB', 'CA4', 'IL7R', 'MYC', 'SPRY1', 'SOX17', 'PROX1', 'CCL21')

FeaturePlot(tiss10x.endo.r2, genes_to_check, pt.size = 2, nCol = 2, cols.use = my.cols, dark.theme = TRUE)

```

```{r, fig.width=10, fig.height=10}
tiss10x.endo.r2.markers <- FindAllMarkers(tiss10x.endo.r2, min.pct = 0.25, thresh.use = 0.25, only.pos = T)

```


```{r, fig.height=80, fig.width=10}
genes_to_check <- tiss10x.endo.r2.markers %>% group_by(cluster) %>% top_n(-4, p_val_adj)
tiss10x.endo.r2.markers %>% group_by(cluster) %>% top_n(-10, p_val_adj)
FeaturePlot(tiss10x.endo.r2, genes_to_check$gene, pt.size = 1, nCol = 2, cols.use = my.cols)
```

## Assigning cell type identity to clusters


```{r, fig.width=10, fig.height=10}
# stash current cluster IDs
tiss10x.endo.r2 <- StashIdent(object = tiss10x.endo.r2, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- 0:max(as.numeric(levels(tiss10x.endo.r2@ident)))
free_annotation <- c('Vein',
                     'Vein',
                     'Capillary Aerocyte',
                     'Vein',
                     'Bronchial Vessel 1',
                     'Vein',
                     'Vein')

tiss10x.endo.r2@meta.data[,'free_annotation'] <- NA
tiss10x.endo.r2@meta.data[,'free_annotation'] <- as.character(plyr::mapvalues(x = tiss10x.endo.r2@ident, from = cluster.ids, to = free_annotation))
```



# Subclustering normal immune cells

```{r, fig.width=10, fig.height=10}
tiss10x.immune = SubsetData(tiss10x, ident.use = c(0,7,16,11,1,9,8,12,18,21,22,19))
```

```{r, fig.width=10, fig.height=10}
tiss10x.immune <- tiss10x.immune %>% ScaleData() %>%
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE, pcs.compute = 30) %>% ProjectPCA(do.print = FALSE)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x.immune, pc.use = 1:9, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.immune, pc.use = 10:18, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.immune, pc.use = 19:27, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)

PCElbowPlot(tiss10x.immune, num.pc = 30)
```


```{r, fig.width=10, fig.height=10}
normal.pcs = c(1,2,3,4,5,6,9,10,13,16,17)
normal.res.use = 1
tiss10x.immune <- tiss10x.immune %>% FindClusters(reduction.type = "pca", dims.use = normal.pcs,
    resolution = normal.res.use, print.output = 0, save.SNN = TRUE, plot.SNN = TRUE, force.recalc = TRUE, prune.SNN = 1/10) %>%
    RunTSNE(dims.use = normal.pcs, seed.use = 10, perplexity=30)

TSNEPlot(object = tiss10x.immune, do.label = T, pt.size = 1, label.size = 10)

```

```{r, fig.width=10, fig.height=10}
TSNEPlot(object = tiss10x.immune, do.return = TRUE, group.by = "orig.ident", label.size = 10)
```

```{r, fig.height=120, fig.width=10}

genes_to_check = c('PTPRC', 'PECAM1', 'EPCAM', 'COL1A1', 'CD3E', 'CD4', 'CD8A', 'GZMH', 'GZMB', 'GZMK', 'CCL5', 'GZMA', 'CD40LG', 'LDHB', 'LEF1', 'CCR7', 'KLRB1', 'TYROBP', 'NKG7', 'FCER1G', 'CD79A', 'CD19', 'MS4A1', 'CD24', 'JSRP1', 'JCHAIN', 'FCGR3B', 'MMP9', 'S100A8', 'MS4A2', 'CPA3', 'RGS13', 'SIGLEC8', 'MARCO', 'MSR1', 'MRC1', 'C1QA', 'LILRB4', 'IRF8', 'LRRC26', 'CLEC4C', 'SMPD3', 'GPR183', 'HLA-DPB1', 'CLEC9A', 'LAMP3', 'CD1C', 'PKIB', 'IGSF21', 'F13A1', 'EREG', 'CLEC5A', 'CHIT1', 'TREM2', 'CD14', 'VCAN', 'FCGR3A', 'CD300E', 'MKI67', 'TOP2A', 'GP9')

FeaturePlot(tiss10x.immune, genes_to_check, pt.size = 1, nCol = 2, cols.use = my.cols, dark.theme = TRUE)

```

```{r, fig.width=10, fig.height=10}
tiss10x.immune.markers <- FindAllMarkers(tiss10x.immune, min.pct = 0.25, thresh.use = 0.25, only.pos = T)

```


```{r, fig.height=500, fig.width=10}
genes_to_check <- tiss10x.immune.markers %>% group_by(cluster) %>% top_n(-4, p_val_adj)
tiss10x.immune.markers %>% group_by(cluster) %>% top_n(-10, p_val_adj)
FeaturePlot(tiss10x.immune, genes_to_check$gene, pt.size = 1, nCol = 2, cols.use = my.cols)
```

## Assigning cell type identity to clusters


```{r, fig.width=10, fig.height=10}
# stash current cluster IDs
tiss10x.immune <- StashIdent(object = tiss10x.immune, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- 0:max(as.numeric(levels(tiss10x.immune@ident)))
free_annotation <- c("Macrophage",
                     "Macrophage",
                     "Macrophage",
                     "Macrophage",
                     "Macrophage",
                     "Macrophage",
                     "Macrophage",
                     NA,
                     NA,
                     NA,
                     "Macrophage",
                     NA,
                     "Macrophage",
                     "Plasma",
                     NA,
                     NA,
                     NA,
                     NA,
                     "Macrophage",
                     "Proliferating Macrophage")

tiss10x.immune@meta.data[,'free_annotation'] <- NA
tiss10x.immune@meta.data[,'free_annotation'] <- as.character(plyr::mapvalues(x = tiss10x.immune@ident, from = cluster.ids, to = free_annotation))
```

# Subclustering normal immune cells (Round 2)

```{r, fig.width=10, fig.height=10}
tiss10x.immune.r2 = SubsetData(tiss10x.immune, ident.use = c(15,17))
```

```{r, fig.width=10, fig.height=10}
tiss10x.immune.r2 <- tiss10x.immune.r2 %>% ScaleData() %>%
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE, pcs.compute = 30) %>% ProjectPCA(do.print = FALSE)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x.immune.r2, pc.use = 1:9, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.immune.r2, pc.use = 10:18, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.immune.r2, pc.use = 19:27, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)

PCElbowPlot(tiss10x.immune.r2, num.pc = 30)
```


```{r, fig.width=10, fig.height=10}
normal.n.pcs = 5
normal.res.use = 1
tiss10x.immune.r2 <- tiss10x.immune.r2 %>% FindClusters(reduction.type = "pca", dims.use = 1:normal.n.pcs,
    resolution = normal.res.use, print.output = 0, save.SNN = TRUE, force.recalc = TRUE, k.param = 10, plot.SNN = TRUE) %>%
    RunTSNE(dims.use = 1:normal.n.pcs, seed.use = 10, perplexity=20)

TSNEPlot(object = tiss10x.immune.r2, do.label = T, pt.size = 3, label.size = 10)

```

```{r, fig.width=10, fig.height=10}
# Batch and sample effects
TSNEPlot(object = tiss10x.immune.r2, do.return = TRUE, group.by = "orig.ident")
```


```{r, fig.height=120, fig.width=10}

genes_to_check = c('PTPRC', 'PECAM1', 'EPCAM', 'COL1A1', 'CD3E', 'CD4', 'CD8A', 'GZMH', 'GZMB', 'GZMK', 'CCL5', 'GZMA', 'CD40LG', 'LDHB', 'LEF1', 'CCR7', 'KLRB1', 'TYROBP', 'NKG7', 'FCER1G', 'CD79A', 'CD19', 'MS4A1', 'CD24', 'JSRP1', 'JCHAIN', 'FCGR3B', 'MMP9', 'S100A8', 'MS4A2', 'CPA3', 'RGS13', 'SIGLEC8', 'MARCO', 'MSR1', 'MRC1', 'C1QA', 'LILRB4', 'IRF8', 'LRRC26', 'CLEC4C', 'SMPD3', 'GPR183', 'HLA-DPB1', 'CLEC9A', 'LAMP3', 'CD1C', 'PKIB', 'IGSF21', 'F13A1', 'EREG', 'CLEC5A', 'CHIT1', 'TREM2', 'CD14', 'VCAN', 'FCGR3A', 'CD300E', 'MKI67', 'TOP2A', 'GP9')

FeaturePlot(tiss10x.immune.r2, genes_to_check, pt.size = 3, nCol = 2, cols.use = my.cols, dark.theme = TRUE)

```

```{r, fig.width=10, fig.height=10}
tiss10x.immune.r2.markers <- FindAllMarkers(tiss10x.immune.r2, min.pct = 0.25, thresh.use = 0.25, only.pos = T, max.cells.per.ident = 200)

```


```{r, fig.height=60, fig.width=10}
genes_to_check <- tiss10x.immune.r2.markers %>% group_by(cluster) %>% top_n(-4, p_val_adj)
tiss10x.immune.r2.markers %>% group_by(cluster) %>% top_n(-10, p_val_adj)
FeaturePlot(tiss10x.immune.r2, genes_to_check$gene, pt.size = 3, nCol = 2, cols.use = my.cols, dark.theme = TRUE)
```

## Assigning cell type identity to clusters

```{r, fig.width=10, fig.height=10}
# stash current cluster IDs
tiss10x.immune.r2 <- StashIdent(object = tiss10x.immune.r2, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- 0:max(as.numeric(levels(tiss10x.immune.r2@ident)))
free_annotation <- c("Memory/Effector T",
                     "Memory/Effector T",
                     "Memory/Effector T",
                     "Basophil/Mast 1",
                     "B",
                     "Naive T",
                     "Natural Killer")

tiss10x.immune.r2@meta.data[,'free_annotation'] <- NA
tiss10x.immune.r2@meta.data[,'free_annotation'] <- as.character(plyr::mapvalues(x = tiss10x.immune.r2@ident, from = cluster.ids, to = free_annotation))
```


# Subclustering normal immune cells (Round 3)

```{r, fig.width=10, fig.height=10}
tiss10x.immune.r3 = SubsetData(tiss10x.immune, ident.use = c(7,8,16))
```

```{r, fig.width=10, fig.height=10}
tiss10x.immune.r3 <- tiss10x.immune.r3 %>%
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE, pcs.compute = 30) %>% ProjectPCA(do.print = FALSE)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x.immune.r3, pc.use = 1:9, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.immune.r3, pc.use = 10:18, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.immune.r3, pc.use = 19:27, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)

PCElbowPlot(tiss10x.immune.r3, num.pc = 30)
```


```{r, fig.width=10, fig.height=10}
normal.pcs = c(1,2,3,5,6,7,8,9)
normal.res.use = 1
tiss10x.immune.r3 <- tiss10x.immune.r3 %>% FindClusters(reduction.type = "pca", dims.use = normal.pcs,
    resolution = normal.res.use, print.output = 0, save.SNN = TRUE, force.recalc = TRUE, plot.SNN = TRUE, k.param = 15, prune.SNN = 1/10) %>%
    RunTSNE(dims.use = normal.pcs, seed.use = 10, perplexity=30)

TSNEPlot(object = tiss10x.immune.r3, do.label = T, pt.size = 3, label.size = 10)

```

```{r, fig.width=10, fig.height=10}
TSNEPlot(object = tiss10x.immune.r3, do.return = TRUE, group.by = "orig.ident", no.legend = T, pt.size = 3)
```

```{r, fig.height=120, fig.width=10}
genes_to_check = c('PTPRC', 'PECAM1', 'EPCAM', 'COL1A1', 'CD3E', 'CD4', 'CD8A', 'GZMH', 'GZMB', 'GZMK', 'CCL5', 'GZMA', 'CD40LG', 'LDHB', 'LEF1', 'CCR7', 'KLRB1', 'TYROBP', 'NKG7', 'FCER1G', 'CD79A', 'CD19', 'MS4A1', 'CD24', 'JSRP1', 'JCHAIN', 'FCGR3B', 'MMP9', 'S100A8', 'MS4A2', 'CPA3', 'RGS13', 'SIGLEC8', 'MARCO', 'MSR1', 'MRC1', 'C1QA', 'LILRB4', 'IRF8', 'LRRC26', 'CLEC4C', 'SMPD3', 'GPR183', 'HLA-DPB1', 'CLEC9A', 'LAMP3', 'CD1C', 'PKIB', 'IGSF21', 'F13A1', 'EREG', 'CLEC5A', 'CHIT1', 'TREM2', 'CD14', 'VCAN', 'FCGR3A', 'CD300E', 'MKI67', 'TOP2A', 'GP9')

FeaturePlot(tiss10x.immune.r3, genes_to_check, pt.size = 2, nCol = 2, cols.use = my.cols, dark.theme = TRUE)
```

```{r, fig.width=10, fig.height=10}
tiss10x.immune.r3.markers <- FindAllMarkers(tiss10x.immune.r3, min.pct = 0.25, thresh.use = 0.25, only.pos = T, max.cells.per.ident = 200)

```


```{r, fig.height=200, fig.width=10}
genes_to_check <- tiss10x.immune.r3.markers %>% group_by(cluster) %>% top_n(-4, p_val_adj)
tiss10x.immune.r3.markers %>% group_by(cluster) %>% top_n(-10, p_val_adj)
FeaturePlot(tiss10x.immune.r3, genes_to_check$gene, pt.size = 1, nCol = 2, cols.use = my.cols)
```

## Assigning cell type identity to clusters

```{r, fig.width=10, fig.height=10}
# stash current cluster IDs
tiss10x.immune.r3 <- StashIdent(object = tiss10x.immune.r3, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- 0:max(as.numeric(levels(tiss10x.immune.r3@ident)))
free_annotation <- c("Classical Monocyte",
                     "IGSF21+ Dendritic",
                     NA,
                     "Myeloid Dendritic Type 2",
                     "Macrophage",
                     "Macrophage",
                     "Classical Monocyte",
                     "Intermediate Monocyte",
                     "Classical Monocyte",
                     "Intermediate Monocyte",
                     "TREM2+ Dendritic",
                     "Proliferating Myeloid Immune",
                     "Classical Monocyte",
                     "TREM2+ Dendritic")


tiss10x.immune.r3@meta.data[,'free_annotation'] <- NA
tiss10x.immune.r3@meta.data[,'free_annotation'] <- as.character(plyr::mapvalues(x = tiss10x.immune.r3@ident, from = cluster.ids, to = free_annotation))
```

# Combine all the annotations into the main object

```{r, fig.width=10, fig.height=10}
tiss10x@meta.data[,'free_annotation'] <- NA

tiss10x@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.epi@meta.data)),'free_annotation'] <- tiss10x.epi@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.epi@meta.data)),'free_annotation']

tiss10x@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.epi.r2@meta.data)),'free_annotation'] <- tiss10x.epi.r2@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.epi.r2@meta.data)),'free_annotation']

tiss10x@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.endo@meta.data)),'free_annotation'] <- tiss10x.endo@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.endo@meta.data)),'free_annotation']

tiss10x@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.endo.r2@meta.data)),'free_annotation'] <- tiss10x.endo.r2@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.endo.r2@meta.data)),'free_annotation']

tiss10x@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.stromal@meta.data)),'free_annotation'] <- tiss10x.stromal@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.stromal@meta.data)),'free_annotation']

tiss10x@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.immune@meta.data)),'free_annotation'] <- tiss10x.immune@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.immune@meta.data)),'free_annotation']

tiss10x@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.immune.r2@meta.data)),'free_annotation'] <- tiss10x.immune.r2@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.immune.r2@meta.data)),'free_annotation']

tiss10x@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.immune.r3@meta.data)),'free_annotation'] <- tiss10x.immune.r3@meta.data[intersect(rownames(tiss10x@meta.data[which(is.na(tiss10x@meta.data[,'free_annotation'])),]),rownames(tiss10x.immune.r3@meta.data)),'free_annotation']

```


```{r, fig.width=10, fig.height=10}
TSNEPlot(tiss10x, do.label = T, group.by = 'free_annotation', no.legend = T)

```

```{r, fig.width=10, fig.height=10}
sort(table(tiss10x@meta.data[,'free_annotation']), decreasing = T)
```

# Removing suspected doublets/cells with low UMIs

```{r, fig.width=10, fig.height=10}
tiss10x.anno = SubsetData(tiss10x, cells.use = rownames(tiss10x@meta.data[which(!is.na(tiss10x@meta.data[,'free_annotation'])),]))
```

```{r, fig.width=10, fig.height=10}
tiss10x.anno <- tiss10x.anno %>% ScaleData() %>%
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE, pcs.compute = 40) %>% ProjectPCA(do.print = FALSE)
```

```{r, fig.width=10, fig.height=10}
PCHeatmap(object = tiss10x.anno, pc.use = 1:9, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.anno, pc.use = 10:18, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.anno, pc.use = 19:27, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)
PCHeatmap(object = tiss10x.anno, pc.use = 28:36, cells.use = 100, do.balanced = TRUE, label.columns = FALSE, num.genes = 20, col.use = my.cols)


PCElbowPlot(tiss10x.anno, num.pc = 40)
```


```{r, fig.width=10, fig.height=10}
n.pcs = 25
tiss10x.anno <- tiss10x.anno %>%
    RunTSNE(dims.use = 1:n.pcs, seed.use = 10, perplexity=30)

TSNEPlot(object = tiss10x.anno, do.label = T, pt.size = 2, label.size = 10, group.by = 'free_annotation', no.legend = TRUE)

```

```{r, fig.width=10, fig.height=10}
# Batch and sample effects
TSNEPlot(object = tiss10x.anno, do.return = TRUE, group.by = "orig.ident", no.legend = T, label.size = 10)

```

# Calculate and plot egression signature of immune cells

```{r}
tiss10x.anno <- ScaleData(tiss10x.anno, genes.use = c('CREM', 'RGS2', 'SLA', 'NFE2L2'))
tmp <- colSums(tiss10x.anno@scale.data[c('CREM', 'RGS2', 'SLA', 'NFE2L2'),])
tiss10x.anno <- AddMetaData(tiss10x.anno, col.name = 'gen.egress.score', tmp)

tiss10x.anno <- ScaleData(tiss10x.anno, genes.use = c('CD69', 'RGS1', 'LMNA', 'RGCC', 'DUSP6', 'SOCS1'))
tmp <- colSums(tiss10x.anno@scale.data[c('CD69', 'RGS1', 'LMNA', 'RGCC', 'DUSP6', 'SOCS1'),])
tiss10x.anno <- AddMetaData(tiss10x.anno, col.name = 'lymphocyte.egress.score', tmp)

tiss10x.anno <- ScaleData(tiss10x.anno, genes.use = c('AREG', 'THBD', 'MPHOSPH6', 'PLAUR', 'HBEGF', 'GNAI3', 'IL1B'))
tmp <- colSums(tiss10x.anno@scale.data[c('AREG', 'THBD', 'MPHOSPH6', 'PLAUR', 'HBEGF', 'GNAI3', 'IL1B'),])
tiss10x.anno <- AddMetaData(tiss10x.anno, col.name = 'myeloid.egress.score', tmp)

# Split groups based on sampling location


VlnPlot(tiss10x.anno, 'gen.egress.score', x.lab.rot = TRUE, group.by = "free_annotation", point.size.use = 0.1) + geom_boxplot()
VlnPlot(tiss10x.anno, 'lymphocyte.egress.score', x.lab.rot = TRUE, group.by = "free_annotation", point.size.use = 0.1) + geom_boxplot()
VlnPlot(tiss10x.anno, 'myeloid.egress.score', x.lab.rot = TRUE,  group.by = "free_annotation", point.size.use = 0.1) + geom_boxplot()
```

# Save the object

```{r, fig.width=10, fig.height=10}
save(tiss10x.anno, file = here('seurat', 'droplet_reyfman_normal_lung_seurat_tiss10x.anno.20200122.RC4.Robj'))
```

```{r, fig.width=10, fig.height=10}
# To reload a saved object
load(file = here('seurat', 'droplet_reyfman_normal_lung_seurat_tiss10x.anno.20200122.RC4.Robj'))
```
